Which ITG Stepcharts are Turniest?

### 2017-03-14 itg/2/paper.tex
Write about turniness.

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}

@@

\keywords
in, the, groove, 2

@@

Right, let's resume right where I left off in my last paper \cite{turniness}, shown in Figure~\ref{fig:you-stutid-fuckass}.
Unlike mainstream conferences, SIGBOVIK doesn't make me waste space repeating all the background material,
and I can just say go read that paper first and get back to me.
It's probably a lot funnier than this one anyway, which is gonna be sort of dry,
and really of interest only to other ITG players who already know what's going on.

@@

@@

\section{Flashback Scene}

Recall Table~\ref{tab:facing} from the last paper, in which I left undefined the facings for $LL$, $DD$, $UU$, and $RR$, the four footswitch footings.

\begin{table}[h]
	\begin{center}
	\begin{tabular}{cc|cccc}
		& & \multicolumn{4}{c}{Right foot} \\
		& & $\leftarrow$ & $\downarrow$ & $\uparrow$ & $\rightarrow$ \\
		\hline
		\multirow{4}{*}{Left foot}
		%                 L   D   U   R
		& $\leftarrow$  & - & $UR$ & $UL$ & $U$ \\
		& $\downarrow$  & $DL$* & - & $L$ & $UL$ \\
		& $\uparrow$    & $DR$* & $R$ & - & $UR$ \\
		& $\rightarrow$ & $D\dagger$ & $DR$* & $DL$* & - \\

	\end{tabular}
	\end{center}
	\caption{Facing directions. ``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
	\label{tab:facing}
\end{table}

@@

% TODO: Put a * on charts where you're going to put a pict of mario man
% TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended
% TODO: Make a special mention of sandstorm and put a pict of mario man

>>>>

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}
\renewcommand{\UrlBreaks}{\do\/\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t\do\u\do\v\do\w\do\x\do\y\do\z}

@@

\keywords
crossovers, footswitches, jacks, sidefoots

@@

Let's resume right where I left off in my last paper \cite{turniness}, shown in Figure~\ref{fig:you-stutid-fuckass}.
Unlike mainstream conferences, SIGBOVIK doesn't make me waste space repeating all the background material,
and I can just say go read that paper first and get back to me.
It's probably a lot funnier than this one anyway, which is gonna be sort of dry,
and really of interest only to other ITG players who already know what's going on.

@@

The TL;DR is that I made a program which figures out how to foot stepcharts in the least crossovery possible way (short of double-stepping everything),
then found which charts ultimately had the most.
The algorithm also naturally identifies footswitches and jacks, %, jacks, forced double-steps, and crossover-footswitches;
and sometimes it's smarter than me in amusing ways.
I put all the goodies in a giant spreadsheet at \url{http://tinyurl.com/crossoveriest},
and the program itself is of course freely available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.

@@

\section{Revisiting Turniness (Flashback Scene)}

Recall Table~\ref{tab:facing} from the last paper, in which I left undefined the facings for $LL$, $DD$, $UU$, and $RR$, the four footswitches.
I show a typical $DD$/$UU$ footswitch pattern in Figure~\ref{fig:fs}(a), and typical $LL$/$RR$ switches (henceforth ``crossover footswitches'') in Figure~\ref{fig:fs}(b).
To step these patterns, the player still alternates feet as usual, but must lift one foot off the repeated arrow before stepping it with her other foot.
Chart authors will often, but not always, include a ``mine cue'' (shown in the figure) to hint that the second foot should switch onto the same arrow.

\begin{table}[t]
	\begin{center}
	\begin{tabular}{cc|cccc}
		& & \multicolumn{4}{c}{Right foot} \\
		& & $\leftarrow$ & $\downarrow$ & $\uparrow$ & $\rightarrow$ \\
		\hline
		\multirow{4}{*}{Left foot}
		%                 L   D   U   R
		& $\leftarrow$  & \bf ? & $UR$ & $UL$ & $U$ \\
		& $\downarrow$  & $DL$ & \bf ? & $L$ & $UL$ \\
		& $\uparrow$    & $DR$ & $R$ & \bf ? & $UR$ \\
		& $\rightarrow$ & $D$ & $DR$ & $DL$ & \bf ? \\

	\end{tabular}
	\end{center}
	\caption{Facing directions.
	%``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
	}
	\label{tab:facing}
\end{table}

\begin{figure}[t]
	\begin{tabular}{c}
	\includegraphics[width=0.13\textwidth]{fs-dd-uu.png} \\
	(a) DD/UU
	\end{tabular}
	\begin{tabular}{c}
	\includegraphics[width=0.13\textwidth]{fs-ll-rr.png} \\
	(b) LL/RR
	\end{tabular}
	\begin{tabular}{c}
	\includegraphics[width=0.13\textwidth]{fs-ll-rr-2.png} \\
	(c) RR/LL
	\end{tabular}
	\caption{Footswitches of various crossoveriness/facing.}
	\label{fig:fs}
\end{figure}

It is tempting to assign the facings $L$, $U$, $U$, and $R$ respectively to the $LL$, $DD$, $UU$, and $RR$ footswitches.
However, Figure~\ref{fig:fs}(c) shows that if a footswitch begins with a crossover on $U$, the facing should be reversed:
the $RR$ footing should face $L$, and $LL$ should face $R$.
``Backwards footswitches'' with $D$ facing are also theoretically possible,
arising from patterns such as $LURDDL$ or $LDRUUL$, although I know of no real-world chart which does this (hmm, tempting...).

Before I realized that, I modified the turniness algorithm \cite{turniness} to face footswitches as above, %the bad way mentioned above,
%to be faced $L$ for $LL$ and so on, as above,
and it surprised me with charts of $\mathcal{T}>2$, in excess of the theoretical maximum!
I show one such chart in Figure~\ref{fig:fs-turniness}(a), in which the step from $LL$ ($\phi=L$) to $UL$ ($\phi=DR$) has individual $\mathcal{T}=3$, and so on for $UL \rightsquigarrow_R UU \rightsquigarrow_L RU \rightsquigarrow_R RR$.
The steps $RR \rightsquigarrow_L LR \rightsquigarrow_R LL$ are both candles ($\mathcal{T}=2$),
resulting overall in $\mathcal{T}=8/3$ for the whole chart.

Indeed, when I further modified the algorithm to force $DD$ switches to face $D$ (i.e., always facing the direction of the repeated arrow),
it produced the chart shown in Figure~\ref{fig:fs-turniness}(b), with overall $\mathcal{T}=3$.
(Note its resemblance to the basic spin pattern, $LDRU$, whose $\mathcal{T}=2$.)

To fairly represent a human player's desire to step in the least turny of ambiguous ways,
I extended the algorithm to provide either the assigned facing, from above, or its polar opposite,
chosen at runtime by whichever is closer to the presequent facing.
This restores the maximum overall chart turniness to $\mathcal{T}=2$,
a new example of which is shown in Figure~\ref{fig:fs-turniness}(c).

However, note that individual steps may still have $\mathcal{T}=3$,
%because after a footswitch's facing is determined, the next step may still ... ?
as shown in Figure~\ref{fig:fs-turniness}(d).
In this example, the step $DL \rightsquigarrow_L LL$ assigns $LL$ to face $L$,
but the subsequent step to $LD$ cannot avoid facing $UR$.
The reason charts still cannot exceed overall $\mathcal{T}=2$ is that
setting up such a situation requires a $\mathcal{T}=1$ step,
which negates the benefit.
A chart could conceivably end right before such a step,
sneaking through some small $\epsilon$ extra turniness \cite{epsilon}
(similar to the case of 270s in \cite{turniness}),
but {\em sustained} average $\mathcal{T}>2$ remains impossible.

Another approach could assign such a footswitch the opposite footing of the previous facing,
%regardless of the switch arrow itself;
so in this case the $LL$ would face $UL$, and each step would have exactly $\mathcal{T}=2$.
Anyway, remember that all-candle switch stream from (c). We'll see it again soon.

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{c}
	\begin{tabular}{c}
	\includegraphics[width=0.13\textwidth]{fs-turniness-8-3.png} \\
	(a) $\mathcal{T}=8/3$(?)%\stackrel{?}{=}8/3$
	\end{tabular}
	\begin{tabular}{c}
	\includegraphics[width=0.13\textwidth]{fs-turniness-3.png} \\
	(b) $\mathcal{T}=3$(?)%\stackrel{?}{=}3$
	\end{tabular}
	\\
	\\
	\begin{tabular}{c}
	\includegraphics[width=0.13\textwidth]{fs-turniness-2.png} \\
	(c) $\mathcal{T}=2$.
	\end{tabular}
	\begin{tabular}{c}
	\includegraphics[width=0.13\textwidth]{fs-turniness-2-plus-epsilon.png} \\
	(d) $\mathcal{T}=2+\epsilon$.
	\end{tabular}
	\end{tabular}
	\end{center}
	\caption{The turniest footswitch patterns. (a) and (b) are false positives (see prose), while (c) and (d) provide theoretically maximal turniness.}
	\label{fig:fs-turniness}
\end{figure}

@@

% TODO: Put a * on charts where you're going to put a pict of mario man
% TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended

% TODO: Make a special mention of sandstorm and put a pict of mario man
% TODO: put the above in an honourable mention subsection, and also mention the slow train and sidefoots
